---
title: "Homework 4"
author: "Anna Giczewska"
format: html
editor: visual
---

## Task 1: Conceptual Questions

``` markdown
1. What is the purpose of the lapply() function? What is the equivalent purrr function??
>  The lapply() function in R is used to apply a function to each element of a list (or vector) and return a list of the same length. The equivalent function in the purrr package is map(). The map() function applies a function to each element of a list (or vector) and returns a list.

2. Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns are numeric). We want use lapply() to run the code cor(numeric_matrix, method = "kendall") on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify method = "kendall" when calling lapply())?
> lapply(X = my_list, FUN = cor, method = "kendall") 

3. What are two advantages of using purrr functions instead of the BaseR apply family?
> The advantage of purrr is much greater consistency between the functions. Purrr also fills in some handy map variants that are absent from base R.

4. What is a side-effect function?
> Side-effect function is a function that doesn't change the data it just try to produce something, like for example: print(), plot(), write_csv(). 

5. Why can you name a variable sd in a function and not cause any issues with the sd function?
> When you call a function, it creates temporary function environments. This is why variables in functions don't exist outside the function (in a global envionment)
```

## Task 2 - Writing R Functions

### Root Mean Square Error (RMSE)

Create some response values and predictions:

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
resp
pred <- predict(lm(resp ~ x), data.frame(x))
pred
```

Write a function:

```{r}
getRMSE <- function(response, prediction, ...) {
  if (length(response) != length(prediction)) {
    stop("The length of the response and prediction vectors is not the same.")
  }
  mse <- mean((response - prediction)^2, ...)
  rmse <- sqrt(mse)
  
  return(rmse)
}
```

Test my function:

```{r}
# Calculate RMSE
rmse_value <- getRMSE(resp, pred)
print(rmse_value)
```

Add NA values:

```{r}
resp[c(3, 10)] <- NA_real_
resp
```

Test function with NAs:

```{r}
# Attempt to calculate RMSE without specifying NA removal
rmse_with_na <- getRMSE(resp, pred)
print(rmse_with_na)

# Calculate RMSE with NA removal
rmse_with_na_rm <- getRMSE(resp, pred, na.rm = TRUE)
print(rmse_with_na_rm)

```

### Mean Absolute Deviation (MAD)

Create some response values and predictions

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

```

Write a function:

```{r}
getMAD <- function(response, prediction, ...) {
  if (length(response) != length(prediction)) {
    stop("The length of the response and prediction vectors is not the same.")
  }
  mad <- mean(abs(response - prediction), ...)
  return(mad)
}
```

Test my function:

```{r}
# Calculate MAD without NA values
mad_value <- getMAD(resp, pred)
print(mad_value)
```

Add NA values:

```{r}
resp[c(3, 10)] <- NA_real_
resp
```

Test function with NAs:

```{r}
# Calculate MAD without specifying NA removal
mad_with_na <- getMAD(resp, pred)
print(mad_with_na)

# Calculate MAD with NA removal
mad_with_na_rm <- getMAD(resp, pred, na.rm = TRUE)
print(mad_with_na_rm)
```

### My wrapper

Create my wrapper: 

```{r}
# Wrapper function
mywrapper <- function(response, prediction, metrics = c("RMSE", "MAD"), ...) {
  # Check that both inputs are numeric, atomic vectors
  if (!(is.vector(response) && is.atomic(response) && is.numeric(response))) {
    stop("Response must be a numeric, atomic vector.")
  }
  if (!(is.vector(prediction) && is.atomic(prediction) && is.numeric(prediction))) {
    stop("Prediction must be a numeric, atomic vector.")
  }
  
  results <- list()
  
  # Calculate requested metrics
  if ("RMSE" %in% metrics) {
    results$RMSE <- getRMSE(response, prediction, ...)
  }
  
  if ("MAD" %in% metrics) {
    results$MAD <- getMAD(response, prediction, ...)
  }
  
  return(results)
}

```

Test new function using data generated below. Call it once asking for each metric individually and once specifying both metrics.

```{r}
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))


mywrapper(resp, pred, metrics = c("RMSE"))
mywrapper(resp, pred, metrics = c("MAD"))
mywrapper(resp, pred, metrics = c("RMSE", "MAD"))

```

Repeat with replacing two of the response values with missing values (NA_real_).

```{r}
resp[c(3, 10)] <- NA_real_
resp

# Calculate both RMSE and MAD with NA removal
both_rmse_mad_na_rm <- mywrapper(resp, pred, na.rm = TRUE)
print(both_rmse_mad_na_rm)

both_rmse_mad<- mywrapper(resp, pred)
print(both_rmse_mad)
  
# Calculate only RMSE
only_rmse_na_rm <- mywrapper(resp, pred, metrics = "RMSE", na.rm = TRUE)
print(only_rmse_na_rm)

only_rmse<- mywrapper(resp, pred, metrics = "RMSE")
print(only_rmse)

# Calculate only MAD
only_mad_na_rm <- mywrapper(resp, pred, metrics = "MAD", na.rm = TRUE)
print(only_mad_na_rm)

only_mad <- mywrapper(resp, pred, metrics = "MAD")
print(only_mad)
```

Finally, test your function by passing it incorrect data (i.e. a data frame or something else instead of vectors)

```{r}
set.seed(10)

# Testing with incorrect data types
# 1. Passing a data frame instead of vectors
incorrect_data <- data.frame(a = rnorm(10), b = rnorm(10))
tryCatch({
  mywrapper(incorrect_data, pred)
}, error = function(e) {
  print(e)
})

# 2. Passing a list instead of vectors
incorrect_list <- list(a = rnorm(10), b = rnorm(10))
tryCatch({
  mywrapper(incorrect_list, pred)
}, error = function(e) {
  print(e)
})

# 3. Passing a character vector instead of numeric vectors
incorrect_char <- as.character(rnorm(10))
tryCatch({
  mywrapper(incorrect_char, pred)
}, error = function(e) {
  print(e)
})

# 4. Passing a matrix instead of vectors
incorrect_matrix <- matrix(rnorm(20), nrow = 10)
tryCatch({
  mywrapper(incorrect_matrix, pred)
}, error = function(e) {
  print(e)
})

```
## Task 3 - Querying an API and a Tidy-Style Function

```{r}
#documentation: https://newsapi.org/docs/endpoints/everything

#install and load needed packages
#install.packages("httr")
library(httr)
#install.packages("jsonlite")
library(jsonlite)

library(tidyverse)

#load API key
api_key <- source("api_key.R")[[1]]

URL_ids <- "https://newsapi.org/v2/everything"
query_keyword <- "technology" 

response_search <- httr::GET(URL_ids, query = list(q = query_keyword, apiKey = api_key))
#str(response_search)
#response_search

# Check the status of the response
if (status_code(response_search) == 200) {
  data <- content(response_search, "text")
  json_data <- jsonlite::fromJSON(data)
  #print(json_data)
  tech_info <- as_tibble(json_data$articles[, -1]) #since source has id and name I want to get rid of them
  print(tech_info)
} else {
  print(paste("Error:", status_code(response_search)))
}

```


Generate my own function:

```{r}
# Load required libraries
library(httr)
library(jsonlite)
library(tibble)

# Define the function
query_news_api <- function(query_keyword, from_date, api_key) {
  URL_ids <- "https://newsapi.org/v2/everything"
  
  # Make the GET request with the API key and query parameters
  response_search <- httr::GET(
    URL_ids, 
    query = list(q = query_keyword, 
                 from = from_date, 
                 to="2024-06-26",  #assuming until today's date not Sys.Date()
                 apiKey = api_key)
    )
  
  # Check the status of the response
  if (status_code(response_search) == 200) {
    data <- content(response_search, "text")
    json_data <- jsonlite::fromJSON(data)
    
    # Print total results
    print(paste("Total results:", json_data$totalResults))
    
    # Convert the articles data to a tibble and remove the first column (source information)
    tech_info <- as_tibble(json_data$articles[, -1]) 
    return(tech_info)
    
  } else {
    stop(paste("Error:", status_code(response_search)))
  }
}

#needed variables
api_key <- source("api_key.R")[[1]]
query_keyword <- "technology"

# Get the news articles
from_date1 <- "2024-06-15"
tech_info1 <- query_news_api(query_keyword, from_date1, api_key)
print(tech_info1) #Seems like by default, the API returns a maximum of 100 articles per request only.

# Get the news articles
from_date2 <- "2024-06-23"
tech_info2 <- query_news_api(query_keyword, from_date2, api_key)
print(tech_info2) #Seems like by default, the API returns a maximum of 100 articles per request only.

```

